% !TEX root = ../../main.tex
% !TEX spellcheck = en_US 
\section{Communication}
\label{sec:communication}
BATS uses the IntentionWriter to send messages to the teammate, a message consists of an
intention together with an optional reason and ping location.
Each intention and reason has multiple messages and are picked at random and put together to form a 
sentence with the intention first—if no reason is specified it will only use the intention as
the whole message. The random function, however, is not entirely random, it will never pick the last
used intention or reason message.

To not spam the teammate messages with the same intention share a timeout, default set to 30\conf~in-game
seconds; if the bot tries to send another message with the same intention before it has timed out
IntentionWriter will skip the message. The timeout can be overridden by each intention, for example
warning that BATS is under attack has a timeout of 60\conf~in-game seconds.

All messages and timeout settings are read from an ini-file to easily add, change, or remove
existing messages, and to reload messages during gameplay. See appendix
\ref{sec:intention_and_reason_messages} for the full ini-file with all intention and reason
messages.

This simple communication system that picks two parts of the sentence at random and put them
together works relatively well. A better approach would be to create
specific messages for each intention and reason together, i.e. not randomize one from intention
and one from reason, but create around 5 messages for each combination. This will create more
tailored messages and will have better English, as this system has some limitations (described
below). This was not implemented because it required much more sentences and thus time.

\paragraph{Limitations}
Because the intention message could be used with many reasons the intention needs to be generic and
cannot be tailored to the situation. E.g. ``Falling back'', ``I'm retreating'', instead of ``Fleeing
from the enemy'', while all work for the reason ``because the enemy is too strong'' the last
intention will not work for ``because you're retreating.''

Many of the sentences may sound a bit strange and some could sound auto-generated. For example ``I
will attack, you're expanding'' and ``Leeeerooooooooy Jenkins, because I'm
expanding''\footnote{Leroy Jenkins is an Internet joke from World of WarCraft where Leroy was tired
of waiting of his group and ran into several mobs (monsters) killing the entire group. In the
beginning he yelled his name, thus this message.}, but maybe not ``I will attack, while you are
expanding'' or ``Sending out an attack, because I'm going to expand.''
