% !TEX root = ../main.tex
% !TEX spellcheck = en_US
\abstract
\begin{changemargin}{+2cm}{+2cm}
\noindent

\marginpar{A bit torn on the context, any ideas?}
\textbf{Context}.
Communication in team games between human player is important, but has been disregarded in teammate bots for Real-Time Strategy (RTS) games. Control over the teammate bots have existed for a while in other genres than RTS; if the bot behaves accordingly it adds another dimension to the game.

\textbf{Objectives}. We investigate whether players think it is more fun to play with a bot that communicates its intentions and reasons and is controllable by the human player. But also what features are liked, disliked, missed to provide guidelines to future researchers and companies.

\textbf{Methods}. An StarCraft RTS bot with communication and control abilities. The experiment consists of four scenarios, turning on/off communication and control, to conclude what players think is fun.

\textbf{Results}. All testers agreed on communication being important and more fun to play with. Beginners did, however, not like the control feature as they already had enough on their mind, experienced players preferred having some control over the bot.

\textbf{Conclusions}. Communication is an important role in team games, including RTS games. Beginners tend to like less control than experienced players. More work needs to be done how to integrate control so that beginners do not feel overwhelmed and experienced players are restrained by to simple commands.

\par\vspace {1cm}
% 3-4 keywords, maximum 2 of these from the title, starts 1 line below the
% abstract.
\noindent
\textbf{Keywords:} AI, RTS, communicating.
\end{changemargin}
\clearpage
