% !TEX root = ../main.tex
\chapter{Problem description and problem statement}
As has been stated, we have yet to find any work on communication with teammate bots for RTS games, to this point. The problem, or rather our hypothesis, is that we think a teammate bot that conveys its intentions and reasons, and can be somewhat controlled by the player to execute various tactics will give more diversity to the game, opening up for new game experiences.

\section{Aims and Objectives?}
Create a teammate bot that communicates its intentions and reasons to the human player and where the control of the bot can be transfered to the player by letting the player send strategic and tactical commands.
\begin{itemize}
	\item Extend Hagelbäck's BTHAI\cite{bthai} to create a bot that
	\begin{itemize}
		\item Sends messages to the player, conveying its intentions and warns about enemy attacks.
		\item Can receive tactical (attack, drop) and strategic (transition to next build order) commands from the player.
		\item Prioritizes the player. 
	\end{itemize}
	\item Evaluate the bot by having human players playing our bot with and without communication, and controls. Evaluation will occur in two stages; first during the later stages of development—to receive feedback on what features need more tweaking—and last to answer the research questions below.
\end{itemize}

\section{Research Questions}
\renewcommand{\theenumii}{\arabic{enumi}.\arabic{enumii}}
\renewcommand{\labelenumii}{\theenumii.}
\begin{enumerate}
	\item	Which is most fun to play with, communicating bot, controllable bot, both communicating and controllable bot, or none?
	\begin{enumerate}
		\item Which features are most liked?
		\item Which features are most disliked?
		\item What features are missing?
	\end{enumerate}
\end{enumerate}

\paragraph{Hypothesis}
\begin{enumerate}
	\item We think it will be most fun to play with both a communicating and controllable bot, then communicating, controllable bot, and last a bot with none of these features.
	\begin{enumerate}
		\item The most liked features are probably the ability to control the bot, that it communicates its intentions to the player, and that it feels as though its programmed to be a good teammate.
		\item We cannot know, since we are not entirely sure how much time we will have left to tweaking.
		\item Same as above.
	\end{enumerate}
\end{enumerate}

The most liked features will probably be the ability to control the bot and that it communicates its intentions and reasons to the player, mostly because the player wants the ability to control the player as he wants it to play and that the player wants to know what the bot is doing. Second to that will be that it feels as though it is programmed to be a good teammate. Disliked features will probably include minor bugs that will not get solved in time, other than that we are not certain how much feature we will have the time to implement; thus we cannot speculate which features the players are going to miss either.

\section{Expected Outcomes}
Game developers can use this teammate bot as a tool to create new fun game experience without having to put as much money into researching whether this would pay off; with an incorrect hypothesis (i.e. not fun) developers know what works and what does not. In general the bot and this paper will provide guidelines how to create, or not create, a teammate bot for RTS games, and what features are most liked, disliked, or missing (i.e. what players might want).

For researchers, this work will either prove that further research in this field will be worthwhile, if the hypothesis is correct; if not, we have at least provided some results what does not work and that this area will not be worthwile to research. In addition for both cases it will provide guidelines for future researchers what to do, and not to do.